Long-duration meeting transcription systems have become increasingly prevalent in both enterprise and professional settings. While speech-to-text accuracy has improved substantially, effective and efficient retrieval of information from lengthy transcripts remains a significant challenge. Meadow9+, a long-meeting transcription application developed by ABAX AI, recently introduced a chatbot feature to address this need; however, its current implementation processes entire transcripts directly as input to a large language model (LLM), resulting in high latency, excessive computational cost, and degraded answer quality due to long-context limitations.

This project proposes and develops a Hybrid Retrieval System to improve transcript retrieval and chatbot performance for Meadow9+. The system combines three coordinated subsystems: a summarization pipeline for high-level transcript understanding, a Retrieval-Augmented Generation (RAG) pipeline for grounded conversational question answering, and a keyword-based search workflow for direct date-based, speaker-based, and lexical retrieval. By building a shared retrieval foundation over transcript chunks and summary-derived representations, the system supports both high-level understanding and fine-grained evidence access, while extending Meadow9+’s transcript retrieval capabilities beyond chatbot-only interaction. 

In addition, this work introduces a comprehensive evaluation pipeline that systematically assesses retrieval quality and end-to-end question-answering performance using metrics such as Hit Rate, Mean Reciprocal Rank (MRR), F1 score, and overall accuracy, alongside detailed latency benchmarking.

Experimental results demonstrate that the RAG-based QA achieves comparable accuracy to the existing full-transcript LLM approach while delivering substantial improvements in response time, validating the effectiveness of retrieval-driven optimization. Beyond immediate chatbot acceleration, the implemented architecture also supports transcript summarization and keyword-based search tailored to noisy speech-to-text outputs. This report presents the design, implementation, and evaluation of the overall Hybrid Retrieval System.
